Scikit-learn: machine learning without learning the machinery

### 2014-05-21 article.tex
Annotate the code
Add forgotten file
Place the code float at the bottom, close to where it is necessary
Compress

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{authblk}
%\usepackage{dblfloatfix}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[numbers,square]{natbib}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{url}
\usepackage{magazine-style}

@@

\paragraph{Project vision}
%
Scikit-learn was born from the observation that most standard
machine-learning algorithms were out of reach of the users that could
most benefit from them: researchers --biologists, climate
scientists, experimental physicists-- or developers --for web
services or domain-specific applications.
%
Implementations mostly consisted of scattered piece of code to download
on researchers' web pages, with the notable exception of
statistic-specific environments such as the R language \cite{Rmanual}, or
the Weka java library \cite{hall2009weka}.

@@

\paragraph{The Python data ecosystem}
%
Machine learning is only a small part of a data-analysis pipeline, and
scikit-learn nicely dovetails into the rich Python ecosystem: scientific
and numeric tools \cite{oliphant2007python,varoquaux2013scipy}, but also
text processing tools, web servers...
%
For its numerical needs, scikit-learn leverages NumPy arrays
\cite{vanderwalt2011} --data structures for efficient numerical
computation--, SciPy --a collection of classic numerical algorithms--,
matplotlib \cite{hunter2007matplotlib} --for scientific plotting-- and
Cython, to generate and call compiled code in a Python-like syntax
\cite{behnel2011cython}.
%
A plethora of Python packages can help users with input or preprocessing
of data, notably Pandas for columnar data \cite{mckinney2012}, scikit-image
for images, and NLTK for text. Finally, the IPython environment
\cite{perez2007ipython} is priceless for interactive work.

\paragraph{Some history}
%
Scikit-learn started circa 2007 as code from David Cournapeau and
Matthieu Brucher's PhD work, but dwindled down till 2010, year at which the
Parietal team from INRIA adopted the project, hiring a full time
engineer. Basics API were defined and an efficient binding of LibSVM
\cite{chang2011libsvm} gave the project a compelling advantage. In
December 2011, the first international sprint was organized with generous
funding from Google. Today, the project has grown vastly beyond
INRIA into a worldwide open source effort.

@@

\paragraph{Setting up models and feeding them data}
%
Models in scikit-learn, called \emph{estimators}, are objects
instantiated with all parameters that describe them, but with data. They
have a {\tt fit} method, that accepts an $(n \times p)$ data matrix,
$\mathbf{X}$ --numpy array or scipy sparse matrix--, where $n$ is the
number of samples, and $p$ the number of features--, and possibly a
quantity to predict, $\mathbf{y}$, also an array, 1D unless the
prediction is \emph{multi-output}, with floats for a regression task, as
in the wages prediction, or integers for classification, that is
predicting for each sample to which class it belongs. The data seldom
come in this form, and they must be adapted to be suitable for
consumption by estimators.

\paragraph{Supervised models: learning to predict}
%
The goal of supervised models is prediction. The corresponding estimators
have a {\tt predict} method, that takes a data matrix $\mathbf{X}$ and
returns a predicted $\mathbf{y}$.

\paragraph{Model evaluation and parameter selection}
%
The data used to train a model can be used to measure the prediction
error, as it would be enable to distinguish noise from signal, and thus
overly optimistic. The correct way is to leave out \emph{test} data,
untouched during training. Often a \emph{cross-validation} scheme is
used, where the data is repeatedly split into \emph{train} and
\emph{test} subsets. Scikit-learn provides a complete framework for
cross-validation. Specific iterators are used to describe the subsets and
several functions and objects accept as arguments these iterators to
perform cross-validation internally. For instance the {\tt
cross\_val\_score} function measures the prediction score of an
estimator. Cross-validation can also be used to tuned the meta-parameters
of an estimator, such as the sparsity of a sparse model. For this,
specific estimators, \emph{meta-estimators}, such as the {\tt
GridSearchCV}, take another estimator at construction time, and use
cross-validation internally to set its parameter. When used together with
{\tt cross\_val\_score}, they perform \emph{nested} cross-validation,
\emph{i.e.}~the meta-parameters are set independently of the test data?

\paragraph{Unsupervised models: learning to transform}
%
Unsupervised learning covers all learning applications in which there is
no clearly identified variable to predict. For instance, in a clustering
application, the task is to group together observations that are similar.
Dimension reduction tries to find simplified representations that capture
well the properties of the data. Novelty detection finds in a new dataset
observations that differ from the data in the train set. As the uses of
unsupervised learning are very diverse, it is not possible to have an API
as uniform as for supervised learning. While some unsupervised
estimators can be used to predict a characteristic of new data, such as
in novelty detection, many are useful to transform data, as in dimension
random. Scikit-learn estimators can have a {\tt transform} method, used
for this purpose. A {\tt Pipeline} estimator can then chain estimators to
form a new estimator that applies transformations before calling the {\tt
fit} or {\tt predict} method of the last object.

\section{In practice}

\paragraph{A simple text-mining example}
%
% Lars: the following text presumes that NumPy, estimators, and the
% X/y convention for supervised learning have already been introduced.

Scikit-learn was designed around NumPy principles and data structures,
so virtually all estimators take arrays of numbers as input
and produce arrays as output.
That doesn't mean that it is limited to traditional number crunching domains,
though.
The library has been successfully employed in text classification
and natural language processing problems,
and utility code is provided to translate the unstructured, symbolic data
encountered in these domains
to the rigid matrix/vector structure that estimators expect.

In the NLP application domain,
features are typically frequencies or boolean indicators
of very large numbers of rare events.
In document classification tasks such spam filtering,
these events may be booleans indicating
``document contains the term \textsf{v1agr@}''
(and so on for all words encountered in the training set).
In other NLP problems such as finding proper names and their classes (NER),
events such as ``next token matches the Roman numeral regex'' may be used.
To ensure scalability in the face of such extremely high-dimensional data,
SciPy's sparse matrix data structures are supported by many estimators.\footnote{
  Support for these is added on a by-need basis,
  because using sparse data structures puts constraints on algorithm design
  and implementation.
}

\begin{figure*}
\lstinputlisting[lastline=12]{nlp_example.py}
\end{figure*}

The following example script is a fully functioning
sentiment polarity classifier for movie reviews:
given a review, it outputs the probability that the review is positive
about the movie it concerns.
The script demonstrates the use of pipelines,
and also shows how machine learning ties into Python's
``batteries included'' philosophy:
downloading, unpacking and learning from a dataset
are all done in one programming language.

% XXX make_pipeline is not in 0.14.1, we should release 0.15 sometime
% The example can be made a bit shorter by skipping the movie_reviews dir;
% the tarball from Pang et al. contain top-level cruft.
%\begin{lstlisting}
%import os.path
%import tarfile
%from urllib import urlretrieve
%from sklearn.datasets import load_files
%from sklearn.feature_extraction.text import TfidfVectorizer
%from sklearn.linear_models import LogisticRegression
%from sklearn.pipeline import make_pipeline
%
%temp, _ = urlretrieve('http://preview.tinyurl.com/review-tgz')
%with tarfile.open(temp) as tar:
%    tar.extractall(path='movie_reviews')
%data = load_files(os.path.join('movie_reviews', 'txt_sentoken'))
%
%clf = make_pipeline(TfidfVectorizer(min_df=2, dtype=float,
%                                    sublinear_tf=True, ngram_range=(1, 2),
%                                    strip_accents='unicode'),
%                     LogisticRegression(random_state=623, C=5000))
%clf.fit(data.data, data.target)
%\end{lstlisting}

This script first fetches the hand-labeled movie review dataset from
\cite{pang2004} and unpacks it to disk using standard Python libraries.
It then loads these into memory using the scikit-learn function
\texttt{load\_files}, makes a pipeline of a \textsf{tf--idf} feature extractor
\cite{rennie2003tackling} and logistic regression,
and finally trains the model.
The feature extractor (vectorizer) takes text as strings
and produces a sparse matrix of weights based on frequencies.
The parameters for the feature extractor and the LogReg classifier
were tuned offline; the random-number generator seed is set by hand
to get reproducible results.

@@

\small
\bibliography{paper}
\bibliographystyle{plain}

>>>>

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{authblk}
\usepackage{dblfloatfix}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[numbers,square]{natbib}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{url}
\usepackage{magazine-style}

@@

\paragraph{Project vision.}
%
Scikit-learn was born from the observation that most standard
machine-learning algorithms were out of reach of the users that could
most benefit from them: researchers --biologists, climate
scientists, experimental physicists-- or developers --for web
services or domain-specific applications.
%
Implementations mostly consisted of scattered piece of code to download
on researchers' web pages, with the notable exception of
statistic-specific environments such as the R language \cite{Rmanual}, or
the Weka java library \cite{hall2009weka}.

@@

\paragraph{The Python data ecosystem.}
%
Machine learning is only a small part of a data-analysis pipeline, and
scikit-learn nicely dovetails into the rich Python ecosystem: scientific
and numeric tools \cite{oliphant2007python,varoquaux2013scipy}, but also
text processing tools, web servers...
%
For its numerical needs, scikit-learn leverages NumPy arrays
\cite{vanderwalt2011} --data structures for efficient numerical
computation--, SciPy --a collection of classic numerical algorithms--,
matplotlib \cite{hunter2007matplotlib} --for scientific plotting-- and
Cython, to generate and call compiled code in a Python-like syntax
\cite{behnel2011cython}.
%
A plethora of Python packages can help users with input or preprocessing
of data, notably Pandas for columnar data \cite{mckinney2012}, scikit-image
for images, and NLTK for text. Finally, the IPython environment
\cite{perez2007ipython} is priceless for interactive work.

\paragraph{Some history.}
%
Scikit-learn started circa 2007 as code from David Cournapeau and
Matthieu Brucher's PhD work, but dwindled down till 2010, year at which the
Parietal team from INRIA adopted the project, hiring a full time
engineer. Basics API were defined and an efficient binding of LibSVM
\cite{chang2011libsvm} gave the project a compelling advantage. In
December 2011, the first international sprint was organized with generous
funding from Google. Today, the project has grown vastly beyond
INRIA into a worldwide open source effort.

@@

\paragraph{Setting up models and feeding them data.}
%
Models in scikit-learn, called \emph{estimators}, are objects
instantiated with all parameters that describe them, but with data. They
have a {\tt fit} method, that accepts an $(n \times p)$ data matrix,
$\mathbf{X}$ --numpy array or scipy sparse matrix--, where $n$ is the
number of samples, and $p$ the number of features--, and possibly a
quantity to predict, $\mathbf{y}$, also an array, 1D unless the
prediction is \emph{multi-output}, with floats for a regression task, as
in the wages prediction, or integers for classification, that is
predicting for each sample to which class it belongs. The data seldom
come in this form, and they must be adapted to be suitable for
consumption by estimators.

\paragraph{Supervised models: learning to predict.}
%
The goal of supervised models is prediction. The corresponding estimators
have a {\tt predict} method, that takes a data matrix $\mathbf{X}$ and
returns a predicted $\mathbf{y}$.

\begin{figure*}[b]
% To have the float on the bottom, we need this declared on the page
% before where it is going to be printed
\lstinputlisting[lastline=12]{nlp_example.py}
\caption{Simple text-processing code to download movie reviews from
Internet and learn positive versus negative ratings.\label{fig:code}}
\end{figure*}

\paragraph{Model evaluation and parameter selection.}
%
The data used to train a model can be used to measure the prediction
error, as it would be enable to distinguish noise from signal, and thus
overly optimistic. The correct way is to leave out \emph{test} data,
untouched during training. Often a \emph{cross-validation} scheme is
used, where the data is repeatedly split into \emph{train} and
\emph{test} subsets. Scikit-learn provides a complete framework for
cross-validation. Specific iterators are used to describe the subsets and
several functions and objects accept as arguments these iterators to
perform cross-validation internally. For instance the {\tt
cross\_val\_score} function measures the prediction score of an
estimator. Cross-validation can also be used to tuned the meta-parameters
of an estimator, such as the sparsity of a sparse model. For this,
specific estimators, \emph{meta-estimators}, such as the {\tt
GridSearchCV}, take another estimator at construction time, and use
cross-validation internally to set its parameter. When used together with
{\tt cross\_val\_score}, they perform \emph{nested} cross-validation,
\emph{i.e.}~the meta-parameters are set independently of the test data?

\paragraph{Unsupervised models: learning to transform.}
%
Unsupervised learning covers all learning applications in which there is
no clearly identified variable to predict. For instance, in a clustering
application, the task is to group together observations that are similar.
Dimension reduction tries to find simplified representations that capture
well the properties of the data. Novelty detection finds in a new dataset
observations that differ from the data in the train set. As the uses of
unsupervised learning are very diverse, it is not possible to have an API
as uniform as for supervised learning. While some unsupervised
estimators can be used to predict a characteristic of new data, such as
in novelty detection, many are useful to transform data, as in dimension
random. Scikit-learn estimators can have a {\tt transform} method, used
for this purpose. A {\tt Pipeline} estimator can then chain estimators to
form a new estimator that applies transformations before calling the {\tt
fit} or {\tt predict} method of the last object.

\section{In practice: putting scikit-learn to good use}

\paragraph{A simple text-mining example.}
%
Not every dataset comes as a set of numbers. For instance in natural
language processing (NLP), each observation is a text, \emph{i.e.} a
string, that needs to be transformed into the data matrix expected by
estimators. Typical, features used to build decisions indicate the
presence of certain words, and often their frequencies.
In spam filtering, it can very relevant to extract from a document that
it contains the term \textsf{viagra}, and possibly how many times.
In other NLP problems such as finding proper names,
events such as ``next token is the Roman numeral'' may be relevant, as in
``J.P. Kennedy III''. Counting bigrams, or $n$-grams, can be a way of
capturing this information. Reusing the {\tt transform} API, scikit-learn
provide a few \emph{feature extraction} objects, that can turn such
unstructured data into a data matrix.
The $\textbf{X}$ matrix is in these situations a
matrix of counting statistics. It is full of zeros, in the sens that very
often a word is present only in a few documents. For better scalability,
it can be encoded using SciPy's sparse matrices, that many scikit-learn
estimators process efficiently.

On Fig.\,\ref{fig:code} we give an example of fully functioning code to
do sentiment polarity classification for movie reviews: given a review,
it outputs whether the review is positive about the movie it concerns.
The code demonstrates the use of pipelines, and also shows how machine
learning ties into unrelated Python's modules: downloading, unpacking and
learning from a dataset are all done in one programming language.

This script first fetches the hand-labeled movie review dataset from
\cite{pang2004} and unpacks it to disk using standard Python libraries.
It then loads these using scikit-learn data loading functions, makes a
pipeline of a \textsf{tf--idf} feature extractor
\cite{rennie2003tackling} --than turns strings into a sparse matrix based
on word frequencies-- and logistic regression, and finally trains the
model. The parameters for the feature extractor and the LogReg classifier
were set manually.

@@

\bibliography{paper}
\bibliographystyle{plain}
